---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scipub

<!-- badges: start -->
<!-- badges: end -->

This package contains functions for summarizing data
    for scientific publication. This includes making a "Table 1"
    to summarize demographics across groups, correlation tables
    with significance indicated by stars, and extracting formatted
    statistical summarizes from simple tests for in-text notation.
    The package also includes functions for Winsorizing data based
    on a Z-statistic cutoff.  
    
Functions:  
    apastat - Format simple statistic test results for scientific publication  
    correltable - Create correlation table (with stars for significance)  for scientific publication  
    FullTable1 - Create Table1 of group summary with stats for scientific publication  
    winsorZ_find - Identify outliers based on z-score cutoff that are Winsorized by the `winsorZ` function  
    winsorZ - Winsorize outliers based on z-score cutoff to next most extreme non-outlier value  

## Installation

You can install the released version of scipub from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("scipub")
```

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("dpagliaccio/scipub")
```
## Example

This is a basic example:

```{r example}
library(scipub)
correltable(data=psydat, html=TRUE)
```


